Reverse Physics: From Laws to Physical Assumptions

### 2022-07-03 Articles/2021-ReversePhysics/ReversePhysics.tex
ReversePhysics: Changing the abstract
Reverse Physics: final changes to archive version

\title{Reverse Physics}
\author{Gabriele Carcassi, Christine A. Aidala}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}

@@

%\begin{itemize}
%	\item Brief descriptions of the aims of foundations of mathematics
%	\item We want to follow a similar approach (not new interpretations of QM or new models for quantum gravity)
%	\item Reverse mathematics start from theorems and finds the smallest set of axioms needed to rederive them. We want to take physical theories and find the smalles set of physical assumptions that are equivalent to the theories themselves.
%	\item This paper is meant to showcase the approach and how it can lead to renewed physical insights into theories that may be thought as already fully understood and those that are not. The main point of the paper is not the results per se, but the novel of the approach proposed. We hope that other people will see the value and pursue similar lines of investigations.
%	
%\end{itemize}

@@

%Use "investigation trumps rigor"
%Somewhere, note that "physics" entropy does not coincide with "information" entropy. Physics entropy is information about states as quantified by invariants.

@@

%\begin{itemize}
%	\item Start with the equations of HM
%	\item Show Liouville's theorem and wonder what it means. Jacobian is 1. Show how Jacobian for 1 d.o.f. is really a Poisson bracket.
%	\item Connect to statistical mechanics: volume is number of states. HM conserves the number of states. States are mapped one to one on a continuum. (discuss problems of counting on a continuum)
%	\item Show that we can go backwards on a single degree of freedom
%	\item Wonder what to do on multiple d.o.f. 
%	\item Show symplectic form as the geometrical description. Generalize to d.o.f. by taking outer product. Show that areas remain orthogonal
%	\item What does orthogonality mean? Independence of d.o.f. (volume product of areas)
%	\item Show that preservation of the geometry (areas and perpendicularity) under time evolution gives us HM
%	\item HM is det/rev motion of indepent d.o.f.
%	\item We used count of states, log count of states is also conserved. What about information entropy? It is also conserved. 
%	\item Show we can proceed the other way. Det/rev means information before and after is the same. Information needs to be conserved -> Jacobian = 1 -> HM.
%	\item Note how bijection is not enough. There is an additional constraint. Why?
%	\item Measurements on a continuum have uncertainty. Connect to statistics and use the covariance matrix to describe a gaussian peak. The determinant of the matrix with the Jacobian. If jacobian is 1, determinant is the same, same precision of the description. HM <-> to requiring a bijection at the same level of precision.
%	\item Wonder why we are getting conservation of energy from det/rev, while we typically associate conservation of energy with system isolation (in the thermodynamic sense). Argue that if a system is deterministic and reversible, it means that the future/past state depends only on the system itself. The system does not depend on anything else -> it is isolated. There is a deep connection between isolation and det/rev
%	
%\end{itemize}

@@

%\begin{itemize}
%	\item We saw that classical mechanics is more than just points: it says something about uncertainty as well. Quantum mechanics has an uncertainty principle. Is there any connections?
%	\item Since there was also a connection to entropy, turn to entropy in quantum mechanics. Entropy is bounded from below: all pure states have zero entropy. In classical mechanics, however, we can prepare distributions with arbitrarily small entropy. Can the fact that the entropy is bound from below in QM be the source of the uncertainty?
%	\item Assume entropy is bounded from below in CM. Show you get uncertainty relationship in a single degree of freedom.
%	\item Uncertainty relationship <-> bounded entropy = bounded information
%\end{itemize}

@@

%\begin{itemize}
%	\item We saw that in quantum mechanics entropy is bounded from below. Why?
%	\item Turn to thermodynamics. The third also says that thermodynamic entropy (not information entropy) is bounded from below. The zero thermodynamic entropy, however, is formulated in terms of a pure crystal at zero temperature. This does not seem to be a fundamental characterization. Can we do better?
%	\item Note how entropy is an extensive quantity. Now consider the empty system. Since the composition of any system with the empty system must yield the same system, it also must yield the same entropy. -> Entropy of empty system must be zero. Better more fundamental characterization of zero entropy.
%	\item What would a system with negative entropy correspond to? To a description of a system that is more precise than saying the system is not there. But that is absurd! Therefore, it is a logical necessity that the entropy of a system is bound from below.	
%\end{itemize}

@@

%\section{A more unified physics}
%
%\begin{itemize}
%	\item Put physics back at the center of the discussion (we are exploring physical concepts and their relationships, the math is there to carry out the arguments)
%	\item Appeal to tradition (thinking about small set of physical assumptions was what foundations of classical mechanics, thermodynamics, ... was about)
%	\item Conceptual unification vs theoretical unification
%	\item Common standard tools (like set theory, category theory, ...)
%	\item Connection between fields
%	\item Reduce the number of fundamental concepts
%	\item These more powerful tools may help theoretical unification
%	\item realm of validity of the different theoretical theories
%	\item realm of validity of the underlying mathematical tools
%\end{itemize}

@@

>>>>

\title{Reverse Physics: From Laws to Physical Assumptions}
\author{Gabriele Carcassi, Christine A. Aidala}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}

@@

@@

@@

@@

@@

@@

@@

\section*{Acknowledgments}
We acknowledge funding from the MCubed program of the University of Michigan. This work is part of a larger project, Assumptions of Physics~\cite{aop-book}, which aims to identify a handful of physical principles from which the basic laws can be rigorously derived.
